---
title: "Dictionary methods"
author: Pablo Barbera
date: May 19, 2016
output: html_document
---


#### Dictionary methods

A different type of keyword analysis consists on the application of dictionary methods, or lexicon-based approaches to the measurement of tone or the prediction of diferent categories related to the content of the text. 

The most common application is sentiment analysis: using a dictionary of positive and negative words, we compute a sentiment score for each individual document.

Let's apply this technique in the context of our running example.

```{r}
library(quanteda)
tweets <- read.csv('../datasets/candidate-tweets.csv', stringsAsFactors=F)
```

```{r}
# loading lexicon of positive and negative words (from Neal Caren)
lexicon <- read.csv("../datasets/lexicon.csv", stringsAsFactors=F)
pos.words <- lexicon$word[lexicon$polarity=="positive"]
neg.words <- lexicon$word[lexicon$polarity=="negative"]
# a look at a random sample of positive and negative words
sample(pos.words, 10)
sample(neg.words, 10)
```

As earlier today, we will convert our text to a corpus object.

```{r}
#install.packages("quanteda")
library(quanteda)
twcorpus <- corpus(tweets$text)
```

Now we're ready to run the sentiment analysis!

```{r}
# first we construct a dictionary object
mydict <- dictionary(list(negative = neg.words,
                          positive = pos.words))
# apply it to our corpus
sent <- dfm(twcorpus, dictionary = mydict)
# and add it as a new variable
tweets$score <- as.numeric(sent[,2]) - as.numeric(sent[,1])
```

```{r}
# what is the average sentiment score?
mean(tweets$score)
# what is the most positive and most negative tweet?
tweets[which.max(tweets$score),]
tweets[which.min(tweets$score),]
# what is the proportion of positive, neutral, and negative tweets?
tweets$sentiment <- "neutral"
tweets$sentiment[tweets$score<0] <- "negative"
tweets$sentiment[tweets$score>0] <- "positive"
table(tweets$sentiment)
```

We can also disaggregate by groups of tweets, for example according to the party they mention.

```{r}
parties <- c("labour", "ukip", "conservative", "snp")

for (p in parties){
  message(p, " -- average sentiment: ",
      round(mean(tweets$score[grep(p, tweets$text, ignore.case=TRUE)]), 4)
    )
}

```

