---
title: "Challenge 1"
author: Pablo Barbera
date: May 19, 2016
output: html_document
---

#### Unsupervised machine learning

Write code in R that help you answer the following questions. We're going to be working with a dataset that contains the abstract of a sample of theses defended at the SPS department at the EUI. We're going to try to see what we can learn from this corpus. 

1. First, read the dataset and create the DFM object.

```{r}
theses <- read.csv('../datasets/eui-theses-data.csv', stringsAsFactors=F)
theses <- theses[!is.na(theses$abstract),]


```

2. Now, estimate the topic model using a number of topics that you think is reasonable.
```{r}

```

3. Look at the words associated with each topic, and a few representative documents for each topic. What do you learn?

```{r}

```

4. Try to find a topic related to terrorism and plots its evolution over time. Has it increased in the past few years?

```{r}

```

BONUS. Experiment with different number of topics. Which value appears to be more appropriate?

```{r}

```