---
title: "Scraping data from Facebook"
author: "Pablo Barbera"
date: "March, 30, 2016"
output: html_document
---

### Scraping web data from Facebook

To scrape data from Facebook's API, we'll use the `Rfacebook` package.

```{r}
#install.packages("Rfacebook")
library(Rfacebook)
```

To get access to the Facebook API, you need an OAuth code. You can get yours going to the following URL: [https://developers.facebook.com/tools/explorer](https://developers.facebook.com/tools/explorer)

Once you're there:  
1. Click on "Get Access Token"  
2. Copy the long code ("Access Token") and paste it here below, substituting the fake one I wrote:

```{r,eval=FALSE}
fb_oauth = 'YOUR_TOKEN_HERE'
```

Now try running the following line:
```{r}
getUsers("me", token=fb_oauth, private_info=TRUE)
```

Does it return your Facebook public information? Yes? Then we're ready to go. See also `?fbOAuth` for information on how to get a long-lived OAuth token.

At the moment, the only information that can be scraped from Facebook is the content of public pages. 

The following line downloads the ~200 most recent posts on the facebook page of the International Criminal Tribunal for the former Yugoslavia
```{r}
page <- getPage("ICTYMKSJ", token=fb_oauth, n=200) 
```

What information is available for each of these posts?
```{r}
page[1,]
```

Which post got more likes, more comments, and more shares?
```{r}
page[which.max(page$likes_count),]
page[which.max(page$comments_count),]
page[which.max(page$shares_count),]
```

We can also subset by date. For example, imagine we want to get all the posts from March 2016 on this Facebook page
```{r}
page <- getPage("ICTYMKSJ", token=fb_oauth, n=1000,
	since='2016/03/01', until='2016/03/30')
page[which.max(page$likes_count),]
```

And if we need to, we can also extract the specific comments from each post.
```{r}
post_id <- page$id[which.max(page$likes_count)]
post <- getPost(post_id, token=fb_oauth, n.comments=1000, likes=FALSE)
```

This is how you can view those comments:
```{r}
comments <- post$comments
head(comments)
```

Also, note that users can like comments! What is the comment that got the most likes?
```{r}
comments[which.max(comments$likes_count),]
```

